Simulating the Linux page cache in distributed applications with the WRENCH library

### 2020-12-08 paper_io_sim.tex
Update section III, IV, figures

            In writeback, one of the keys is the cache eviction mechanism, which is
            triggered when available memory is insufficient to removes pages 
            from page cache. 
            Because dirty pages are data written to page cache but yet to be 
            written to disk, dirty pages must be kept in page cache and only clean pages 
            are available for eviction.
            If the clean pages are inadequate, dirty pages are flushed to disk, 
            marked as clean and available for eviction. 
            There is another type of flushing mechanism, which is 
            \textit{periodical flushing}, called periodically in predefined intervals. 
            Periodical flushing only flushes expired dirty pages, which remain dirty in 
            page cache longer than an expiration time configured in the kernel. 
            Different cache eviction algorithms have also been developed and proposed
            \cite{owda2014comparison}.
            In Linux kernel, a \textit{two lists strategy (LRU/2)} is implemented based on 
            the least-recently-used (LRU) scheme, with an \textit{active list} and 
            an \textit{inactive list}. 
            If accessed pages are not in the page cache, they are added to the inactive list. 
            If pages are accessed when residing in the inactive list, they will be moved from 
            inactive list to active list. 
            The lists are also kept balanced by moving pages from the active list
            to the inactive list when the active list grows too large.
            Thus, the active list only contains pages which are accessed more than once 
            and not evictable, while the inactive list includes pages accessed once only, 
            or pages accessed more than once but moved from the active list. 
            The lists work in LRU manner, pages are added to the tail and removed 
            from the top.

@@

    We separate our simulation model in two components, the IO
    Controller and the Memory Manager, which together simulate 
    file reads and writes in Linux (Figure~\ref{fig:interaction}). 
    To read or write a file chunk, a simulated application sends a
    request to the IO Controller. The IO Controller interacts as needed with
    the Memory Manager to free memory through flushing or eviction,
    and to read or write to cache. The Memory Manager
    implements these operations, simulates periodical flushing
    and eviction from cache, and reads or writes to disk when necessary.
    \begin{figure}
           \centering
           \includegraphics[width=0.85\columnwidth]{figures/interaction.pdf}
           \caption{Overview of the simulated writeback page cache \tristan{does it simulate only writeback? See comment in NFS section.}.
           Applications send file read or write requests to the
           IO Controller that orchestrates flushing, eviction, cache
           and disk accesses with the Memory Manager. Concurrent accesses to storage
           devices (memory and disk) are simulated using existing models.}
           \label{fig:interaction}
    \end{figure}

@@

    We model page cache LRU lists as 
    two lists of data blocks, an active list and an inactive list, both ordered by 
    last access time (earliest first, see Figure~\ref{fig:lrulist}).
    As in the kernel, our simulator limits the size of the active list to
    twice the size of the inactive list, by moving least recently 
    used data blocks from the active list to the inactive list~\cite{gorman2004understanding, linuxdev3rd2010}.
    % Modeling page cache LRU lists as lists of data blocks reduces the
    % overhead of the simulator, as data blocks are obviously less
    % numerous than file pages. At the same time, it achieves the accuracy in 
    % I/O time simulation that can be seen in the later section.
    % From Tristan: I don't think the sentences above are necessary,
    % they are quite repetitive.

    At any given time, a file can be partially cached, completely cached,
    or uncached at all. Cached data blocks may reside in one or both of the
    page cache LRU lists. The first time they are accessed, blocks are
    added to the inactive list. On subsequent accesses, blocks of the
    inactive list are moved to \tristan{if they are moved then they can't
    be in both lists} the top of the active list. Cached blocks written to
    cache are marked as dirty until they are flushed to disk.

@@

    Our simulation model supports chunk-by-chunk file accesses
    with a user-defined chunk size. However, for simplicity, we assume that file pages are 
    accessed in a round-robin fashion rather than fully randomly. 
    Therefore, when a file is read, cached data is read only after all uncached data was read, and data from the inactive list is read
    before data from the active list
    (data reads occur from left to right in Figure~\ref{fig:read_order}).
    When a chunk of \emph{uncached} data is read, a new clean block is created 
    and appended to the inactive list. 
    When a chunk of \emph{cached} data is read, one or more existing data blocks in the LRU lists are accessed.
    If these blocks are clean, we merge them together, update the access time and size of the resulting block, 
    and append it to the active list. 
    If the blocks are dirty, we move them independently to the active list, to preserve their entry time. 
    Because the chunk and block sizes may be different, there are situations
    where a block is not entirely read. 
    In this case, the block is split in two smaller blocks and one of them is re-accessed.
    % From Tristan: not sure where to put this nor if it's necessary:
    % , in which chunks are read/written
    % until file is entirely read/written.
    \begin{figure}
           \centering
           \includegraphics[width=\columnwidth]{figures/read_order.pdf}
           \caption{File data read order. Data is read from left to right: uncached data 
           is read first, followed by data from the inactive list, and finally data from the active list. \tristan{Colors 
           shouldn't match that in Fig 1 as they dont represent the same thing.}}
           \label{fig:read_order}
    \end{figure}

@@

    \begin{algorithm}\caption{Periodical flushing simulation}\label{alg:pdflush}
        \small
        \begin{algorithmic}[1]
            \Input
                \Desc{in}{page cache inactive list}
                \Desc{ac}{page cache active list}
                \Desc{t}{predefined flushing time interval}
                \Desc{exp}{predefined expiration time}
                \Desc{sm}{storage simulation model}
               \EndInput
               \While{host is on}
                \State blocks = expired\_blocks(exp, in) + expired\_blocks(exp, ac) 
                \State flushing\_time = sm.write(blocks)
                \State flag blocks as clean data \tristan{a bit informal}
                \If{flushing\_time $<$ t}
                    \State sleep(t - flushing\_time)
                \EndIf \tristan{else?}
            \EndWhile
        \end{algorithmic}
    \end{algorithm}

@@

    \begin{algorithm}\caption{File chunk read simulation of IO Controller}
    \label{alg:read}
        \small
        \begin{algorithmic}[1]
            \Input
                \Desc{cs}{chunk size}
                \Desc{fn}{file name}
                \Desc{fs}{file size (assumed to fit in memory)}
                \Desc{mm}{MemoryManager object}
                \Desc{sm}{storage simulation model}
               \EndInput
               \State disk\_read = min(cs, fs - mm.cached(fn)) \Comment{To be read from disk} 
               \State cache\_read = cs - disk\_read \Comment{To be read from cache} 
            \State mm.flush(cs + disk\_read - mm.free\_mem - mm.evictable, fn) 
            \State mm.evict(cs + disk\_read - mm.free\_mem, fn) 
            \If {disk\_read $>$ 0}  \Comment{Read uncached data}    
                \State sm.read(disk\_read)  
                \State mm.add\_to\_cache(disk\_read, fn)     
            \EndIf
            \If {cache\_read $>$ 0} \Comment{Read cached}
                \State mm.cache\_read(cache\_read)  
            \EndIf
        \end{algorithmic}
    \end{algorithm}
    As mentioned previously, our model reads and writes file chunks in a
    round-robin fashion. To read a file chunk, simulated applications send
    chunk read requests to the IO Controller which processes them using
    Algorithm~\ref{alg:read}. First, we calculate the amount of uncached
    data that needs to be read from disk, and the remaining amount is read
    from cache (line 7-8). The amount of memory required to read the chunk
    is calculated, corresponding to a copy of the chunk in anonymous memory
    and a copy of the chunk in cache \tristan{Create a variable
    "required\_mem" in the algorithm to clarify}.
    \tristan{add a call to mm in the algorithm to increment anonymous memory. Mention in the experiments that anonymous memory is flushed after each task.}
    If there is not enough available memory, the Memory Manager is called
    to flush dirty data (line 9). If necessary, flushing is complemented by
    eviction (line 10). Note that, when called with negative arguments, functions
    \texttt{flush} and \texttt{evict} simply return and do not do anything. Then, if the block requires
    uncached data, a disk read is simulated with the disk model (line 12)
    \tristan{this is not consistent with the new Figure 1 where disk reads
    are done by the mm}, and the data read is added to cache (line
    13).
    Finally, if cached data needs to be read, the Memory Manager is called
    to simulate a cache read  
    and update the corresponding data blocks accordingly (line 16).

    \begin{algorithm}\caption{File chunk write simulation of IO Controller}
    \label{alg:write}
        \small
        \begin{algorithmic}[1]
            \Input
                \Desc{cs}{chunk size}
                \Desc{fn}{file name}
                \Desc{mm}{MemoryManager object}
                \Desc{sm}{storage simulation model}
               \EndInput
            \State remain\_dirty = dirty\_ratio * mm.avail\_mem - mm.dirty
            \If {remain\_dirty $>$ 0} \Comment{Write with memory bandwidth}
                \State mm.evict(min(cs, remain\_dirty) - mm.free\_mem)
                \State mem\_amt = min(cs, mm.free\_mem)
                \State mm.write(fn, mem\_amt) 
            \EndIf
            \If {mem\_amt $<$ cs}  \Comment{Write with disk bandwidth}
                \State mm.flush(cs - mem\_amt)  
                \State mm.evict(cs - mem\_amt  - mm.free\_mem) 
                \State mm.add\_to\_cache(fn, min(cs - mem\_amt, mm.free\_mem)) \tristan{why is data written to cache with "addtocache" here and with "write" at line 10?}
                \State sm.write(cs - mem\_amt) \tristan{According to Fig 1 the IO Controller is not supposed to interact directly with sm}
            \EndIf

        \end{algorithmic}
    \end{algorithm}
    Algorithm~\ref{alg:write} describes our simulation of chunk writes in 
    the IO Controller. 
    % Data is written to cache until the amount of dirty data in the cache 
    % exceeds the dirty ratio times the amount of available memory. Data is then written to disk.
    Our algorithm initially checks the  amount of data that 
    can be written as dirty data given the dirty ratio (line 6).
    If this amount is greater than 0, the Memory Manager is requested to evict 
    data from cache if necessary (line 8) \tristan{the difference between available and free memory should be explained the first time it is referred}.
    After eviction, the amount of data that can be written to 
    page cache is calculated (line 9), and a cache write is simulated (line 10).
    To write the remaining data, we flush and 
    evict data as possible so that this data can be 
    written to cache (line 13-14). 
    The Memory Manager does not guarantee that the amount of flushed 
    and evicted data is enough to write the remaining data. 
    In the case it is less than the required amount, the remaining data 
    being written to cache can be flushed and evicted right away. 
    Thus, the amount of remaining data written and kept in cache 
    after the write is limited by the amount free memory, and this amount 
    is added to cache (line 15) \tristan{the previous two sentences are very unclear, I don't understand them so I can't reformulate.}. 
    Finally, the disk write is simulated with the remaining data amount (line 16), 
    and the chunk write algorithm finishes \tristan{see comment in algorithm}.

@@

            We also implemented our model in WRENCH, separately from existing WRENCH components. 
            We exposed page cache simulation through a user API, activated with a command-line argument.
            We used SimGrid 3.25 and WRENCH 1.6. SimGrid source code is available at \url{https://framagit.org/simgrid/simgrid},
            and WRENCH is at \url{https://github.com/wrench-project/wrench/tree/master/src/wrench/services/memory}.
            \tristan{which commit did you use?}

@@

        We simulate an application that consists of three single-core, sequential tasks
        where each task reads the file produced by the previous task,
        increments every byte of this file to emulate real processing, and
        writes the resulting data to disk. We also use a real application
        \textcolor{red}{[Maybe briefly describe this application's purpose
        (e.g. a standard fMRI preprocessing piepeline)]}, to evaluate the
        applicability of our simulation model.

@@

        To parametrize our simulators, we benchmarked the bandwidths of
        memory, local disk, remote disk (NFS), and network
        (Table~\ref{table:benchmark}). Since WRENCH currently only supports
        symmetrical bandwidths, we use the mean of the read and write
        bandwidth values in our experiments.
            \begin{table}[htbp]
            \centering
            \begin{tabularx}{\columnwidth}{|c
            |>{\centering\arraybackslash}X
            |>{\centering\arraybackslash}X
            |>{\centering\arraybackslash}X|}
            \hline
                Device  & Cluster (real) & Python simulator & WRENCH simulator\\
            \hline
                Memory read  & 6860    & 4812     & 4812\\
                Memory write & 2764    & 4812 & 4812\\
                Local disk read & 510 & 465 & 465\\
                Local disk write & 420 & 465     & 465\\
                Remote disk read & 515 & - & 445\\
                Remote disk write & 375 & - & 445\\
                Network bandwidth & 3000 & - & 3000\\
            \hline
            \end{tabularx}
            \caption{Bandwidth benchmarks (MBps) and simulator configurations.
            The bandwidths used in the simulations are the average of the read and write bandwidths
            measured in the cluster.
            Remote disk and network accesses are not simulated in the Python simulator. \tristan{table could be prettified}}
            \label{table:benchmark}
            \end{table}

            As our focus is on I/O rather than compute, we measured
            application task durations on a cluster node and used these
            durations in our simulation. For the Python simulator,
            execution times were directly used in the simulation. For
            WRENCH and WRENCH-cache, we determined the corresponding number
            of flops on a 1~Gflops CPU and used these values in the
            simulation (see Table~\ref{table:tasks} \tristan{add a table to describe execution times}). The final simulated
            platform and application used in our experiment are avaiable at
            \url{https://github.com/wrench-project/wrench/tree/ec6b43561b95977002258c0fe37a4ecad8f1d33f/examples/basic-examples/io-pagecache}.

@@

            Our third experiment occurs in the same configuration as the
            previous one, albeit reading and writing on a 400-GB remote NFS disk
            mounted from another compute node. \tristan{As is commonly
            configured in HPC environments to avoid data loss,?} NFS client
            and server read caches are enabled, but server writes are configured with
            write-through cache instead of write back \tristan{how about client write
            cache?}. Therefore, all the writes happen at disk bandwidth, but
            reads can benefit from cache hits. We simulate write-through by
            writing data directly to disk and adding them to the server
            cache \tristan{it it part of the IO Controller? it would deserve
            a few sentences in the methods.}.

@@

            \tristan{mention that cache was flushed before each experiment}

@@

        Figure~\ref{fig:single_error} shows the relative error of simulated
        read and write times throughout the three application tasks. The
        page cache simulation model drastically reduces simulation errors
        for all I/O operations except the first read that deals with
        uncached data. Errors are reduced from an average of \texttt{xxx\%}
        in WRENCH to \texttt{yyy \%} in the Python prototype and
        \texttt{zzz \%} in WRENCH-cache. The original WRENCH simulator significantly overestimates
        read and write times, due to its lack of page cache simulation
        model.

        % In this figure, \textit{file $<$i$>$} is the input of \textit{Read $<$i$>$} and
        % output of \textit{Write $<$i-1$>$}.
        With the input size of 20 GB, in both real execution and WRENCH-cache
        simulator, there is no difference in the page cache since all files in the
        pipeline can be entirely cached \tristan{what does this refer to? 
        For 20 GB there is still some error in WRENCH-cache, what do you mean ``there is no difference in the page cache''?}.

        With the input of 100 GB, the difference between real execution and WRENCH-cache
        simulator can be seen in the amount of \textit{file 3} cache after \textit{Write 2}.
        In real execution, after being written by \textit{Write 2},
        \textit{file 3} is entirely in cache.
        Thus, in \textit{Read 3}, \textit{file 3} is read with memory bandwidth.
        However, in the WRENCH-cache simulator, only a part of \textit{file 3} is
        cached after \textit{Write 2}, which means \textit{Write 2} reads a part of
        \textit{file 3} from disk.
        The reason for the difference is in our WRENCH-cache simulator,
        with 100 GB of input, while \textit{file 3} is being written,
        the page cache is saturated, eviction is then evoked to evict data
        from page cache and make space available for \textit{file 3}.
        However, because the evicted amount is less than required,
        some \textit{file 3} data is evicted after being written, make \textit{file 3}
        partially cached.
        As a result, in \textit{Read 3}, only a part of \textit{file 3} is read from cache
        with memory bandwidth, a part of it is read from disk with disk read bandwidth,
        which is many times slower, leading to longer read time.

@@

>>>>

            In writeback, one of the keys is the cache eviction mechanism, which is
            triggered when available memory is insufficient to remove pages 
            from page cache. 
            The term \textit{available memory} here includes both free memory and 
            the memory used but can be reclaimed by the kernel with flushing 
            and eviction. 
            Because dirty pages are data written to page cache but yet to be 
            written to disk, dirty pages must be kept in page cache and only clean pages 
            are available for eviction.
            If the clean pages are inadequate, dirty pages are flushed to disk, 
            marked as clean and available for eviction. 
            There is another type of flushing mechanism, which is 
            \textit{periodical flushing}, called periodically in predefined intervals. 
            Periodical flushing only flushes expired dirty pages, which remain dirty in 
            page cache longer than an expiration time configured in the kernel. 
            Different cache eviction algorithms have also been developed and proposed
            \cite{owda2014comparison}.
            In Linux kernel, a \textit{two lists strategy (LRU/2)} is implemented based on 
            the least-recently-used (LRU) scheme, with an \textit{active list} and 
            an \textit{inactive list}. 
            If accessed pages are not in the page cache, they are added to the inactive list. 
            If pages are accessed when residing in the inactive list, they will be moved from 
            inactive list to active list. 
            The lists are also kept balanced by moving pages from the active list
            to the inactive list when the active list grows too large.
            Thus, the active list only contains pages which are accessed more than once 
            and not evictable, while the inactive list includes pages accessed once only, 
            or pages accessed more than once but moved from the active list. 
            The lists work in LRU manner, pages are added to the tail and removed 
            from the top.

@@

    We separate our simulation model in two components, the IO
    Controller and the Memory Manager, which together simulate 
    file reads and writes in Linux (Figure~\ref{fig:interaction}). 
    To read or write a file chunk, a simulated application sends a
    request to the IO Controller. The IO Controller interacts as needed with
    the Memory Manager to free memory through flushing or eviction,
    and to read or write to cache. The Memory Manager
    implements these operations, simulates periodical flushing
    and eviction from cache, and reads or writes to disk when necessary.  
    In case write-through strategy is used, IO Controller can also interact directly 
    with disk model to simulate disk writes.

    \begin{figure}
           \centering
           \includegraphics[width=0.85\columnwidth]{figures/interaction.pdf}
           \caption{Overview of the page cache simulator.
           Applications send file read or write requests to the
           IO Controller that orchestrates flushing, eviction, cache
           and disk accesses with the Memory Manager. Concurrent accesses to storage
           devices (memory and disk) are simulated using existing models.}
           \label{fig:interaction}
    \end{figure}

@@

    We model page cache LRU lists as 
    two lists of data blocks, an active list and an inactive list, both ordered by 
    last access time (earliest first, see Figure~\ref{fig:lrulist}).
    As in the kernel, our simulator limits the size of the active list to
    twice the size of the inactive list, by moving least recently 
    used data blocks from the active list to the inactive list~\cite{gorman2004understanding, linuxdev3rd2010}.

    At any given time, a file can be partially cached, completely cached,
    or uncached at all. A cached data block can only reside in one of two 
    page cache LRU lists. The first time they are accessed, blocks are
    added to the inactive list. On subsequent accesses, blocks of the
    inactive list are moved to the top of the active list. Cached blocks 
    written to cache are marked as dirty until they are flushed to disk.

@@

    Our simulation model supports chunk-by-chunk file accesses
    with a user-defined chunk size. However, for simplicity, we assume that file pages are 
    accessed in a round-robin fashion rather than fully randomly. 
    Therefore, when a file is read, cached data is read only after all uncached data was read, and data from the inactive list is read
    before data from the active list
    (data reads occur from left to right in Figure~\ref{fig:read_order}).
    When a chunk of \emph{uncached} data is read, a new clean block is created 
    and appended to the inactive list. 
    When a chunk of \emph{cached} data is read, one or more existing data blocks in the LRU lists are accessed.
    If these blocks are clean, we merge them together, update the access time and size of the resulting block, 
    and append it to the active list. 
    If the blocks are dirty, we move them independently to the active list, to preserve their entry time. 
    Because the chunk and block sizes may be different, there are situations
    where a block is not entirely read. 
    In this case, the block is split in two smaller blocks and one of them is re-accessed.
    % From Tristan: not sure where to put this nor if it's necessary:
    % , in which chunks are read/written
    % until file is entirely read/written.
    \begin{figure}
           \centering
           \includegraphics[width=\columnwidth]{figures/read_order.pdf}
           \caption{File data read order. Data is read from left to right: uncached data 
           is read first, followed by data from the inactive list, and finally data from the active list. }
           \label{fig:read_order}
    \end{figure}

@@

    \begin{algorithm}\caption{Periodical flushing simulation}\label{alg:pdflush}
        \small
        \begin{algorithmic}[1]
            \Input
                \Desc{in}{page cache inactive list}
                \Desc{ac}{page cache active list}
                \Desc{t}{predefined flushing time interval}
                \Desc{exp}{predefined expiration time}
                \Desc{sm}{storage simulation model}
               \EndInput
               \While{host is on}
                \State blocks = expired\_blocks(exp, in) + expired\_blocks(exp, ac) 
                \State flushing\_time = sm.write(blocks)
                \State for blk in blocks: blk.dirty = 0
                \If{flushing\_time $<$ t}
                    \State sleep(t - flushing\_time)
                \EndIf  % Else: do nothing, continue the loop
            \EndWhile
        \end{algorithmic}
    \end{algorithm}

@@

    \begin{algorithm}\caption{File chunk read simulation of IO Controller}
    \label{alg:read}
        \small
        \begin{algorithmic}[1]
            \Input
                \Desc{cs}{chunk size}
                \Desc{fn}{file name}
                \Desc{fs}{file size (assumed to fit in memory)}
                \Desc{mm}{MemoryManager object}
                \Desc{sm}{storage simulation model}
               \EndInput
               \State disk\_read = min(cs, fs - mm.cached(fn)) \Comment{To be read from disk}
               \State cache\_read = cs - disk\_read \Comment{To be read from cache}
               \State required\_mem = cs + disk\_read
               \State mm.flush(cs + disk\_read - mm.free\_mem - mm.evictable, fn) 
               \State mm.evict(cs + disk\_read - mm.free\_mem, fn) 
               \If {disk\_read $>$ 0}  \Comment{Read uncached data}    
               \State sm.read(disk\_read)  
               \State mm.add\_to\_cache(disk\_read, fn)     
               \EndIf
               \If {cache\_read $>$ 0} \Comment{Read cached}
               \State mm.cache\_read(cache\_read)  
            \EndIf
            \State mm.use\_anonymous\_mem(cs)
        \end{algorithmic}
    \end{algorithm}
    As mentioned previously, our model reads and writes file chunks in a
    round-robin fashion. To read a file chunk, simulated applications send
    chunk read requests to the IO Controller which processes them using
    Algorithm~\ref{alg:read}. First, we calculate the amount of uncached
    data that needs to be read from disk, and the remaining amount is read
    from cache (line 6-7). The amount of memory required to read the chunk
    is calculated, corresponding to a copy of the chunk in anonymous memory
    and a copy of the chunk in cache (line 8). 
    If there is not enough available memory, the Memory Manager is called
    to flush dirty data (line 9). If necessary, flushing is complemented by
    eviction (line 10). Note that, when called with negative arguments, functions
    \texttt{flush} and \texttt{evict} simply return and do not do anything. Then, if there is
    uncached data, memory manager is called to read data from disk and add this 
    data to cache (line 12).
    If cached data needs to be read, the Memory Manager is called to simulate 
    a cache read  and update the corresponding data blocks accordingly (line 15).
    Finally, the memory manager is called to reduce the amount of free memory 
    corresponding to the amount of anonymous memory used by application (line 17).

    \begin{algorithm}\caption{File chunk write simulation of IO Controller}
    \label{alg:write}
        \small
        \begin{algorithmic}[1]
            \Input
                \Desc{cs}{chunk size}
                \Desc{fn}{file name}
                \Desc{mm}{MemoryManager object}
               \EndInput
            \State remain\_dirty = dirty\_ratio * mm.avail\_mem - mm.dirty
            \If {remain\_dirty $>$ 0} \Comment{Write with memory bandwidth}
                \State mm.evict(min(cs, remain\_dirty) - mm.free\_mem)
                \State mem\_amt = min(cs, mm.free\_mem)
                \State mm.write\_to\_cache(fn, mem\_amt) 
            \EndIf
            \State remaining = cs - mem\_amt
            \While {remaining $>$ 0}  \Comment{Write with disk bandwidth}
                \State mm.flush(cs - mem\_amt)
                \State mm.evict(cs - mem\_amt  - mm.free\_mem) 
                \State to\_cache = min(remaining, mm.free\_mem)
                \State mm.write\_to\_cache(fn, to\_cache) 
                \State remaining = remaining - to\_cache
            \EndWhile

        \end{algorithmic}
    \end{algorithm}
    Algorithm~\ref{alg:write} describes our simulation of chunk writes in 
    the IO Controller. 
    % Data is written to cache until the amount of dirty data in the cache 
    % exceeds the dirty ratio times the amount of available memory. Data is then written to disk.
    Our algorithm initially checks the  amount of data that 
    can be written as dirty data given the dirty ratio (line 5).
    If this amount is greater than 0, the Memory Manager is requested to evict 
    data from cache if necessary (line 7).
    After eviction, the amount of data that can be written to 
    page cache is calculated (line 8), and a cache write is simulated (line 9). 
    If there is remaining data, this data is written to cache in a loop (line 12 - line 18).
    In this loop, we flush and evict data as possible so that there will be 
    available space for writing (line 13-14).  
    After flushing and eviction, the amount of data that can be written 
    is calculated (line 15) and a cache write is simulated (line 16). 
    We update the remaining amount (line 17) and continue the loop until all data 
    is written.

    In addition to the chunk write simulation with writeback page cache above, 
    we also add a chunk write function with write-through strategy to the IOController.
    This function simply simulates a disk write with the amount of data passed in, 
    then evicts cache if needed and adds this data to cache.

@@

            We also implemented our model in WRENCH, separately from existing WRENCH components. 
            We exposed page cache simulation through a user API, activated using a command-line argument.
            We used SimGrid 3.25 which is available at \url{https://framagit.org/simgrid/simgrid}.
            To implement the simulator, we used the WRENCH 1.6 version at this commit \url{https://github.com/wrench-project/wrench/tree/67185374330d2c4bf274fce222c937e838df5b03}.

@@

        We simulate an application that consists of three single-core, sequential tasks
        where each task reads the file produced by the previous task,
        increments every byte of this file to emulate real processing, and
        writes the resulting data to disk. 
        In real execution, we clear cache before starting the application to make sure 
        the conditions are consistent, the cache is not in use and it is only used by 
        our application.
        In addition, the anonymous memory used by application 
        is released after each task in the pipeline. 
        Thus, we also release the memory used at the end of each task in our simulator.
        We also use a real application
        \textcolor{red}{[Maybe briefly describe this application's purpose
        (e.g. a standard fMRI preprocessing piepeline)]}, to evaluate the
        applicability of our simulation model.

@@

        To parametrize our simulators, we benchmarked the bandwidths of
        memory, local disk, remote disk (NFS), and network
        (Table~\ref{table:benchmark}). Since WRENCH currently only supports
        symmetrical bandwidths, we use the mean of the read and write
        bandwidth values in our experiments.
            \begin{table}[htbp]
            \centering
            \begin{tabularx}{\columnwidth}{|l
            |>{\centering\arraybackslash}X
            |>{\centering\arraybackslash}X
            |>{\centering\arraybackslash}X|}
            \hline
                Bandwidths  & Cluster (real) & Python simulator & WRENCH simulator\\
            \hline
                Memory read  & 6860    & 4812     & 4812\\
                Memory write & 2764    & 4812 & 4812\\
                Local disk read & 510 & 465 & 465\\
                Local disk write & 420 & 465     & 465\\
                Remote disk read & 515 & - & 445\\
                Remote disk write & 375 & - & 445\\
                Network & 3000 & - & 3000\\
            \hline
            \end{tabularx}
            \caption{Bandwidth benchmarks (MBps) and simulator configurations.
            The bandwidths used in the simulations are the average of the read and write bandwidths
            measured in the cluster.
            Remote disk and network accesses are not simulated in the Python simulator. \tristan{table could be prettified}}
            \label{table:benchmark}
            \end{table}

            As our focus is on I/O rather than compute, we measured
            application task durations on a cluster node and used these
            durations in our simulation. For the Python simulator,
            execution times were directly used in the simulation. For
            WRENCH and WRENCH-cache, we determined the corresponding number
            of flops on a 1~Gflops CPU and used these values in the
            simulation (see Table~\ref{table:cputime}). The final simulated
            platform and application used in our experiment are avaiable at
            \url{https://github.com/wrench-project/wrench/tree/ec6b43561b95977002258c0fe37a4ecad8f1d33f/examples/basic-examples/io-pagecache}.

            \begin{table}[htbp]
            \centering
            \begin{tabularx}{0.8\columnwidth}{|l|>{\centering\arraybackslash}X|}
            \hline
                Input size (GB)  & CPU time (second)\\
            \hline
                3 	 & 4.4 \\
                20  & 28 \\
                50  & 75 \\
                75  & 110 \\
                100  & 155 \\
            \hline
            \end{tabularx}
            \caption{Task CPU time corresponding to input file size.}
            \label{table:cputime}
            \end{table}

@@

            Our third experiment occurs in the same configuration as the
            previous one, albeit reading and writing on a 400-GB NFS-mounted remote disk
            mounted from another compute node. As is commonly
            configured in HPC environments to avoid data loss, in our simulator, 
            NFS client and server read caches are enabled, but there is no cache 
            for writing on client side and server writes are configured with
            write-through cache instead of writeback. 
            Therefore, all the writes happen at disk bandwidth, but
            reads can benefit from cache hits. We simulate write-through by
            writing data directly to disk and adding that data to the server
            cache.

@@

@@

@@

             Figure~\ref{fig:single_error} shows the relative error of simulated
             read and write times throughout the three application tasks. The
             page cache simulation model drastically reduces simulation errors
             for all I/O operations except the first read that deals with
             uncached data. Errors are reduced from an average of 345\%
             in WRENCH to 46\% in the Python prototype and 39\% in WRENCH-cache.
             The original WRENCH simulator significantly overestimates
             read and write times, due to its lack of page cache simulation
             model.
             Especially, the simulation time errors of \textit{Read 3} with 100 GB of input
             in the simulators with page cache are just under the error in original WRENCH.
             This longer read can be explained with the amount of cached data of files
             at the beginning of the read in reality and simulators shown in
             Fig.~\ref{fig:single_cache}.
             % In this figure, \textit{file $<$i$>$} is the input of \textit{Read $<$i$>$} and
             % output of \textit{Write $<$i-1$>$}.
             With the input size of 20 GB, the amount of cached data of each files after 
             every I/O operation is similar since the page cache can keep all these files.
%             \tristan{what does this refer to?
%             For 20 GB there is still some error in WRENCH-cache, what do you mean 
%``there is no difference in the page cache''?}.
%   Dzung: Some text was lost in a merge, I added it again.
             With the input of 100 GB, the difference between real execution and WRENCH-cache
             simulator can be seen in the amount of \textit{file 3} cache after \textit{Write 2}.
             In real execution, after being written by \textit{Write 2},
             \textit{file 3} is entirely in cache.
             Thus, in \textit{Read 3}, \textit{file 3} is read with memory bandwidth.
             However, in the WRENCH-cache simulator, only a part of \textit{file 3} is
             cached after \textit{Write 2}, which means \textit{Write 2} reads a part of
             \textit{file 3} from disk.
             The reason for the difference is in our WRENCH-cache simulator,
             with 100 GB of input, while \textit{file 3} is being written,
             the page cache is saturated, eviction is then evoked to evict data
             from page cache and make space available for \textit{file 3}.
             However, because the evicted amount is less than required,
             some \textit{file 3} data is evicted after being written, make \textit{file 3}
             partially cached.
             As a result, in \textit{Read 3}, only a part of \textit{file 3} is read from cache
             with memory bandwidth, a part of it is read from disk with disk read bandwidth,
             which is many times slower, leading to longer read time.
